---
title: "Manuscript Tables and Figures"
author: "Raahil Madhok"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    number_sections: yes
  pdf_document: default
---

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = '/Users/rmadhok/Dropbox (Personal)/ebird_lockdown/data/')
```

```{r,include=F,warning=FALSE,message=FALSE}
# Load Packages
rm(list=ls())
require(tidyverse)
require(lubridate)
require(kableExtra)
require(cowplot)
source('/Users/rmadhok/Dropbox (Personal)/ebird_lockdown/scripts/R/select_functions.R')
```

# Overview
Sample selection protocol:

1. Stationary trips

2. Complete checklists (report all species)

3. Drop duplicates from group trips.

4. Drop trips with 1 bird

5. Duration between 5-240 mins

Overall Notes

1. Results are sensitive to sample selection criteria (num cities, number days pre/post, home, lockdown date)

2. We select cities first, and then users. This gives us the same users in each pre/post period (in 2020) with the caveat that users mobility is limited to the top X cities.  

3. Diff-in-diff results consistently positive and strong. Event study somewhat unstable depending on lockdown day. Why is DID counterfactual better?  

## Birdwatching Activity in All Cities
Summary Stats for 2020 including users who participated in at least 2 days before and after lockdown.

```{r,include=F,warning=FALSE,message=FALSE}
# 2020 sample
ebird_full <- readRDS('ebd_full.rds')

# 24 days before and after lockdown (Mar 24 2020)
ebird <- ebird_full[ebird_full$YEAR == 2020 & 
                      ebird_full$OBSERVATION.DATE <= '2020-04-17',]
```

```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.height=3,fig.width=8}
# Number of Users by day
#stats <- select_sample(ebird)
stats <- ebird %>%
  group_by(OBSERVATION.DATE) %>%
  summarize(`Num. Users` = n_distinct(OBSERVER.ID),
            `Num. Trips` = n_distinct(SAMPLING.EVENT.IDENTIFIER),
            `Num. Species` = n_distinct(TAXONOMIC.ORDER)) %>%
  gather(key, value, starts_with('N'))

ggplot(stats, aes(x=OBSERVATION.DATE, y=value)) +
  geom_line() + 
  labs(title = 'Birdwatching Activity in all Cities', y='') +
  geom_vline(xintercept=as.numeric(as.Date("2020-03-24")), linetype='dashed') +
  scale_x_date(date_breaks = '10 days', date_labels = '%b\n%d') +
  theme(panel.grid.minor = element_blank(), 
        axis.line = element_blank(), 
        axis.title.x = element_blank()) +
  facet_wrap(~key, scales='free')
```

## Birdwatching Activity in Top 4 Cities

```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.height=10,fig.width=8}
# By City ## WHAT IS GOING ON??
#stats <- select_sample(ebird, num_cities = 20)
#stats <- select_cities(ebird, num=20)
stats <- ebird %>%
  group_by(COUNTY, OBSERVATION.DATE) %>%
  summarize(`Num. Users` = n_distinct(OBSERVER.ID),
            `Num. Trips` = n_distinct(SAMPLING.EVENT.IDENTIFIER),
            `Num. Species` = n_distinct(TAXONOMIC.ORDER)) %>%
  gather(key, value, starts_with('N'))
stats <- filter(stats, COUNTY %in% c('Mumbai', 'Bangalore', 'Chennai', 'Kolkata'))

ggplot(stats, aes(x=OBSERVATION.DATE, y=value)) +
  geom_line() +
  labs(title = 'Birdwatching Activity in Four Major Cities', y='') +
  geom_vline(xintercept=as.numeric(as.Date("2020-03-24")), linetype='dashed') +
  scale_x_date(date_breaks = '10 days', date_labels = '%b\n%d') +
  theme(panel.grid.minor = element_blank(), 
        axis.line = element_blank(), 
        axis.title.x = element_blank()) +
  facet_wrap(~COUNTY+key, scales='free', ncol=3)
```

## Top 20 Cities
This is the subset of eBird observations meeting the pariticipation criteria in the top 20 Indian cities by population size. Note that not all top 20 cities may be present in eBird. 

```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.height=6,fig.width=8}
stats <- select_sample(ebird, num_cities = 20)
stats <- stats %>% 
  group_by(COUNTY) %>%
  summarize(STATE = first(STATE), POP_DENSITY = first(POP_DENSITY)) %>%
  arrange(-POP_DENSITY)

kable(stats, 
      caption = 'eBird Cities in Top 20 by Population Density') %>% 
  kable_styling()
```

## Pre-Post Lockdown Statistics
This table shows aggregate birdwatching activity in the full sample among users who birdwatched on at least 2 days before and after lockdown (March 24, 2020). 
```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.height=6,fig.width=8}

# All Cities
stats <- select_sample(ebird)
stats <- stats %>%
  mutate(PREPOST = if_else(
    OBSERVATION.DATE < '2020-03-24', 'Before', 'After')
    ) %>%
  group_by(PREPOST) %>%
  summarize(`Num. Trips` = n_distinct(SAMPLING.EVENT.IDENTIFIER),
            `Num. Cities` = n_distinct(COUNTY),
            `Num. Users` = n_distinct(OBSERVER.ID),
            `Num. Species` = n_distinct(TAXONOMIC.ORDER)) %>%
  arrange(desc(PREPOST))

kable(stats, caption = 'All Cities') %>% kable_styling()

# Top 20
stats <- select_sample(ebird, num_cities = 20)
stats <- stats %>%
  mutate(PREPOST = if_else(
    OBSERVATION.DATE < '2020-03-24', 'Before', 'After')
    ) %>%
  group_by(PREPOST) %>%
  summarize(`Num. Trips` = n_distinct(SAMPLING.EVENT.IDENTIFIER),
            `Num. Cities` = n_distinct(COUNTY),
            `Num. Users` = n_distinct(OBSERVER.ID),
            `Num. Species` = n_distinct(TAXONOMIC.ORDER)) %>%
  arrange(desc(PREPOST))

kable(stats, caption = 'Top 20') %>% kable_styling()
```

## Trip Statistics
This table shows trip-level statistics before and after lockdown.

```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.height=6,fig.width=8}

# Full Sample
stats <- select_sample(ebird)
stats <- stats %>%  
  distinct(SAMPLING.EVENT.IDENTIFIER, .keep_all = T) %>%
  dplyr::select(OBSERVATION.DATE, SAMPLING.EVENT.IDENTIFIER, 
                DURATION.MINUTES, s_richness, COUNTY) %>%
  mutate(PREPOST = if_else(
    OBSERVATION.DATE < '2020-03-24', 'Before', 'After')
    ) %>%
  group_by(PREPOST) %>%
  summarize(`Trip Length` = mean(DURATION.MINUTES),
            `Mean Species Richness` = mean(s_richness, na.rm=T),
            `Median Species Richness` = median(s_richness, na.rm=T)) %>%
  arrange(desc(PREPOST))

kable(stats, caption='All Cities') %>% kable_styling()

# Top 20
stats <- select_sample(ebird, num_cities = 20)
stats <- stats %>%  
  distinct(SAMPLING.EVENT.IDENTIFIER, .keep_all = T) %>%
  dplyr::select(OBSERVATION.DATE, SAMPLING.EVENT.IDENTIFIER, 
                DURATION.MINUTES, s_richness, COUNTY) %>%
  mutate(PREPOST = if_else(
    OBSERVATION.DATE < '2020-03-24', 'Before', 'After')
    ) %>%
  group_by(PREPOST) %>%
  summarize(`Trip Length` = mean(DURATION.MINUTES),
            `Mean Species Richness` = mean(s_richness, na.rm=T),
            `Median Species Richness` = median(s_richness, na.rm=T)) %>%
  arrange(desc(PREPOST))

kable(stats, caption='Top 20') %>% kable_styling()

# By City
stats <- select_sample(ebird, num_cities = 20)
stats <- stats %>%  
  distinct(SAMPLING.EVENT.IDENTIFIER, .keep_all = T) %>%
  dplyr::select(OBSERVER.ID, OBSERVATION.DATE, SAMPLING.EVENT.IDENTIFIER, 
                DURATION.MINUTES, s_richness, COUNTY) %>%
  mutate(PREPOST = if_else(
    OBSERVATION.DATE < '2020-03-24', 'Before', 'After')
    ) %>%
  group_by(COUNTY, PREPOST) %>%
  summarize(`Trip Length` = mean(DURATION.MINUTES),
            `Mean Species Richness` = mean(s_richness, na.rm=T),
            `Median Species Richness` = median(s_richness, na.rm=T)) %>%
  filter(!(COUNTY %in% c('Central Delhi', 'New Delhi')))

ggplot(stats, aes(x=factor(PREPOST), y=`Mean Species Richness`)) + 
  geom_bar(stat = 'identity') + 
  scale_x_discrete(limits=c('Before','After')) +
  labs(title = 'Mean Species Richness Per Trip') +
  theme(panel.grid.minor = element_blank(), 
        axis.line = element_blank(), 
        axis.title.x = element_blank(),
        axis.ticks = element_blank()) +
  facet_wrap(~COUNTY, scales = 'free')
```

## Marginal Species by city
Marginal species are species observed after lockdown which were not observed before lockdown. To implement: in each city, obtain list of species observed (across all users) in before and after lockdown. Extract species in "after" list not in "before" list, and collect in a data frame. If fewer species after, city will not be in final list. 

```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.height=6,fig.width=8}
# Species List
stats <- select_sample(ebird, num_cities = 20)
stats$prepost <- if_else(stats$OBSERVATION.DATE < '2020-03-24', 'PRE', 'POST')
city_list <- distinct(stats, COUNTY)$COUNTY
marginal_species <- data.frame()

# stack Marginal Species per city
for(city in city_list) {
  
  df_pre <- stats %>%
    filter(COUNTY == city & prepost == 'PRE') %>%
    distinct(COMMON.NAME, .keep_all = T) %>%
    select('COMMON.NAME', 'SCIENTIFIC.NAME', 'COUNTY')

  df_post <- stats %>%
    filter(COUNTY == city & prepost == 'POST') %>%
    distinct(COMMON.NAME, .keep_all = T) %>%
    select('COMMON.NAME', 'SCIENTIFIC.NAME', 'COUNTY')
  
  marginal <- anti_join(df_post, df_pre, by='COMMON.NAME')
  marginal_species <- rbind(marginal_species, marginal)
}

kable(marginal_species) %>%
  kable_styling(fixed_thead = T)
rm(list=c('city_list', 'df_post', 'df_pre', 'marginal'))
```

# Event Study (March 24 Lockdown)
We estimate:

$$ SR_{ijtd} = \sum_{e=-24, e\neq0}^{24} 1(t-LD=e)\beta_{e} + X_{ijdt} + \gamma_{t} + \mu_{d} + \epsilon_{ijtd}$$
where $SR_{ijtd}$ is species richness observed by user i on trip j on date t in district d. Each coefficient $\beta_{e}$ captures species richness relative to the day of lockdown (LD). In the preferred specification, LD = March 24, 2020 and this date is normalized to zero and omitted so that it describes the reference period. If the timing of lockdown is exogenous to pre-period outcomes, then $\beta_{e}$ should be zero for all $e<0$. $\gamma_{t}$ are a set of temporal controls and include date fixed effects and hour fixed effects. $\mu_{d}$ are city fixed effects. $X_{ijdt}$ is a vector of controls and includes: duration of trip and user $i$'s experience, measured as the number of trips he/she has taken since January 2019. 

Notes: 

1. We see action around pilot lockdown, not as much around official lockdown. Suggests people responded to the former. 

2. We start seeing increases in species diversity post-lockdown when we narrow in on urban environments. As we increase allowable cities, this adds population-sparse places so the effect is dampned. 

3. Larger increases in post-lockdown species diversity as we increase the participation constraint (e.g. 10 days before/after curfew)


```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.height=15,fig.width=12}
stats <- data.frame()

# All Cities, 2 days before/after
coefs <- event_study(ebird, user_fe=T)
coefs$type <- 'Cities: all, participation: 2 days, LD: Mar 24'
stats <- rbind(stats, coefs)

# Top 20, 2 days before/after
coefs <- event_study(ebird, num_cities = 20, user_fe=T)
coefs$type <- 'Cities: 20, participation: 2 days, LD: Mar 24'
stats <- rbind(stats, coefs)

# Top 20, 5 days before/after
coefs <- event_study(ebird, num_cities = 20, before = 5, after = 5, user_fe=T)
coefs$type <- 'Cities: 20, participation: 5 days, LD: Mar 24'
stats <- rbind(stats, coefs)

# Top 20, 10 days before/after
coefs <- event_study(ebird, num_cities = 20, before = 10, after = 10, user_fe=T)
coefs$type <- 'Cities: 20, participation: 10 days, LD: Mar 24'
stats <- rbind(stats, coefs)

# Top 50, 5 days before/after
coefs <- event_study(ebird, num_cities = 50, before = 5, after = 5, user_fe=T)
coefs$type <- 'Cities: 50, participation: 5 days, LD: Mar 24'
stats <- rbind(stats, coefs)

ggplot(stats, aes(x=time, y=estimate, group=1)) + 
  geom_line() + 
  geom_point(aes(color='Coefficient')) +
  geom_ribbon(aes(ymin=conf.low, ymax=conf.high, fill='95% CI'), alpha=.5) +
  geom_vline(xintercept=0, linetype='dashed') +
  geom_hline(yintercept=0, linetype='dashed') +
  geom_text(aes(x=-2, label="Lockdown", y=2), angle=90) +
  ylab('Species Richness Relative to Lockdown\n') +
  xlab('\n Days to Lockdown') +
  scale_x_continuous(breaks = seq(-20, 30, by = 5)) +
  scale_colour_manual("" , values = "black") +
  scale_fill_manual("", values = "grey12") +
  theme(panel.grid.minor = element_blank(),
        text = element_text(size=13),
        axis.line = element_blank(), 
        axis.ticks = element_blank()) +
  facet_wrap(~type, scales='free', ncol=2)
```

## Species Distributions (Bangalore Only)
data: number of times each user sees each species before and after lockdown.
interpretation: proportion of user base seeing different frequencies of species.

```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.height=100,fig.width=20, unit='in'}
stats <- select_sample(ebird) %>%
  mutate(prepost = if_else(OBSERVATION.DATE < '2020-03-24', 'PRE', 'POST'))

bangalore <- stats %>%
  filter(COUNTY == 'Bangalore') %>%
  fastDummies::dummy_cols(select_columns='COMMON.NAME') %>%
  group_by(OBSERVER.ID, prepost) %>%
  summarize_at(vars(starts_with('COMMON.NAME_')), sum, na.rm=T) %>%
  gather(species, number, starts_with('COMMON.NAME'))

ggplot(data = bangalore, 
       aes(x = number, 
           fill=prepost)) +
  geom_histogram(alpha=0.5, aes(y=..density..), position="identity") +
  facet_wrap(~species, ncol=4)

```

# Difference-in-Difference (PRELIMINARY)

In this specification, we aim to estimate the difference in species richness per trip before and after lockdown, compared to same difference in 2019. In other words, we subtract the change in the pre-post period of 2019 (the counterfactual) from the 2020 lockdown effect to remove other determinants of species richness that are changing during the window. 

To do this, we first select 2020 users with activity in at least 2 days before and after lockdown. Then we add all users from 2019 for the same days. These are different users but share similar characteristics.

Formally, we specify the following standard DID regression:

$$ SR_{ijdt} = \beta_{1} [Treatment_{t} \times Post_{t}] + \beta_{2} [Treatment_{t}] + \beta_{3}[Post_{t}] + X_{ijdt} + \mu_{d} + \gamma_{t} + \epsilon_{ijdt} $$

Where $Treatment$ is a dummy for the year 2020 and $Post$ is a dummy equal to one for observations after lockdown. $\beta_{1}$ is the coefficient of interest and describes the change in species richness before and after March 19 in 2020 compared to the same difference in 2019. $\beta_{2}$ describes the difference in the "before" groups (pre-intervention). $\beta_{3}$ captures the time trend in the control (2019) group. $X_{ijdt}$ is a vector of controls including trip duration. $\mu_{d}$ is a set of city fixed effects and $\gamma_{t}$ is a set of temporal controls including date and hour fixed effects. 

```{r,echo=FALSE,warning=FALSE,message=FALSE,fig.height=6,fig.width=8}
require(modelsummary)
models <- list()

#-----------MARCH 24 Lockdown ------------

# 2019 sample
ebird_19 <- ebird_full[ebird_full$YEAR == 2019 &
                         ebird_full$OBSERVATION.DATE <= '2019-04-17',]
ebird_19 <- select_sample(ebird_19, num_cities=20, lockdown = '2019-03-24')
ebird_19 <- ebird_19 %>%
  distinct(SAMPLING.EVENT.IDENTIFIER, .keep_all = T) %>%
    dplyr::select(OBSERVER.ID, OBSERVATION.DATE, SAMPLING.EVENT.IDENTIFIER,
                  YEAR, DURATION.MINUTES, s_richness, COUNTY, STATE, 
                  HOUR, n_trips_pld) %>%
    mutate(prepost = if_else(OBSERVATION.DATE < '2019-03-24', 'PRE', 'POST'))

# 2020 sample
ebird_20 <- select_sample(ebird, num_cities=20)
ebird_20 <- ebird_20 %>%
  distinct(SAMPLING.EVENT.IDENTIFIER, .keep_all = T) %>%
    dplyr::select(OBSERVER.ID, OBSERVATION.DATE, SAMPLING.EVENT.IDENTIFIER, 
                  YEAR, DURATION.MINUTES, s_richness, COUNTY, STATE, 
                  HOUR, n_trips_pld) %>%
    mutate(prepost = if_else(OBSERVATION.DATE < '2020-03-24', 'PRE', 'POST'))

data <- rbind(ebird_19, ebird_20)
data$Treatment <- as.numeric(data$YEAR == 2020)
data$Post <- as.numeric(data$prepost == 'POST')
data$TreatPost <- data$Treatment*data$Post

models[['Mar 24 Lockdown']] <- lm(s_richness ~ 
                                    TreatPost + Treatment + Post + DURATION.MINUTES +
                                    COUNTY + OBSERVATION.DATE + HOUR, data=data)

## ---------- MARCH 22 Lockdown -------------
ebird_19 <- ebird_full[ebird_full$YEAR == 2019 &
                         ebird_full$OBSERVATION.DATE <= '2019-04-17',]
ebird_19 <- select_sample(ebird_19, num_cities=20, lockdown = '2019-03-22')
ebird_19 <- ebird_19 %>%
  distinct(SAMPLING.EVENT.IDENTIFIER, .keep_all = T) %>%
    dplyr::select(OBSERVER.ID, OBSERVATION.DATE, SAMPLING.EVENT.IDENTIFIER,
                  YEAR, DURATION.MINUTES, s_richness, COUNTY, STATE, 
                  HOUR, n_trips_pld) %>%
    mutate(prepost = if_else(OBSERVATION.DATE < '2019-03-22', 'PRE', 'POST'))

ebird_20 <- select_sample(ebird, num_cities=20, lockdown = '2020-03-22')
ebird_20 <- ebird_20 %>%
  distinct(SAMPLING.EVENT.IDENTIFIER, .keep_all = T) %>%
    dplyr::select(OBSERVER.ID, OBSERVATION.DATE, SAMPLING.EVENT.IDENTIFIER, 
                  YEAR, DURATION.MINUTES, s_richness, COUNTY, STATE, 
                  HOUR, n_trips_pld) %>%
    mutate(prepost = if_else(OBSERVATION.DATE < '2020-03-22', 'PRE', 'POST'))

data <- rbind(ebird_19, ebird_20)
data$Treatment <- as.numeric(data$YEAR == 2020)
data$Post <- as.numeric(data$prepost == 'POST')
data$TreatPost <- data$Treatment*data$Post

models[['Mar 22 Lockdown']] <- lm(s_richness ~ 
                                    TreatPost + Treatment + Post + DURATION.MINUTES +
                                    COUNTY + OBSERVATION.DATE + HOUR, data=data)

cm <- c('TreatPost'= 'Treat x Post',
        'DURATION.MINUTES'='Duration',
        'Post' = 'Post',
        'Treatment'='Treatment')
msummary(models, coef_map=cm, stars=T,
         title = "Diff-in-Diff Estimates (Top 20 cities)")
```
